---
title: "Untitled"
output: html_document
---


---
title: "ch2 4th corner September 2018"
output: html_document
---

<br>
<br>

##Introduction

<br>

Host plant genotype and phenotype influence arthropod community composition. They influence fitness, can mediate competition, predation and host-parasite interactions. Plant size, leaf chemistry, defense compounds, trichome density etc all affect species in different ways. However, it is hard to pin down general principles in how host plant characteristics affect herbivores across species 
Insect traits can be used to provide a mechanistic connection between species and environment. Insect traits have increasingly been used to understand insect response to characteristics of landscape and understand community composition.

<br>

I set out with the following objectives:


##### (i) How do herbivore communities respond to phenotypic variation?
##### (ii) Which plant traits account for herbivore community responses to host-plant phenotype?
##### (iii) Which insect traits account for herbivore community responses to host-plant phenotype



<br>
<br>


```{r Load_data&packages, echo=FALSE, warning=FALSE, message=FALSE}
rm(list=ls())


library(ggplot2)
library(scales)
library(lattice)
library(vegan)
library(iNEXT)
library(plyr)
library(mvabund)
library(reshape2)
library(lme4)

#try species by site dataset, including leps and coleoptera
setwd("/Users/elsketielens/phd/data/HDIM/branch clipping/data/data for R")
hem<-read.table("hem.sep18.csv", header=TRUE, sep="")

##load other datasets 
setwd("/Users/elsketielens/phd/data/HDIM/branch clipping/data/data for R")
env<- read.table("env site july16.csv", header=TRUE, sep=",")
env<-env[env$vialID%in%rownames(hem),]
setwd("/Users/elsketielens/phd/data/HDIM/branch clipping/data/data for R")
traits<-read.table("trait species july16new.csv", header=T, sep=",")
setwd("/Users/elsketielens/phd/data/HDIM/branch clipping/data/data for R")
leaftraits.full<-read.table("leaf traits july16.csv", header=TRUE, sep=",")


# Subset to only four sites, for both years
env.both<-env[env$site!="ERY",]
env.both<-env.both[env.both$site!=c("TPR"),]
env.both<-env.both[env.both$site!=c("KAIO"),]
env.both<-env.both[env.both$site!=c("THU"),]
env.both<-env.both[env.both$site!=c("LH"),]
env.both<-env.both[env.both$site!=c("LY"),]
env.both<-env.both[env.both$site!=c("L65"),]
env.both<-env.both[env.both$site!=c("KOHY"),]
env.both<-env.both[env.both$site!=c("KOHO"),]
env.both<-env.both[env.both$island=="Hawaii",]

# create env data for 2014
env.2014<-env.both[env.both$sampleyear=="2014",]


# create hem data
hem.both<-hem[rownames(hem)%in%env.both$vialID,]
hem.2014<-hem[rownames(hem)%in%env.2014$vialID,]

## Create traits data 
#traits.both<-traits[rownames(traits)%in%colnames(hem.both),]
traits<-traits[traits$spec_code%in%colnames(hem.2014),]

# exclude species that don't occur in our four sites
hem.both<-hem.both[,colSums(hem.both)>0]
hem.2014<-hem.2014[,colSums(hem.2014)>0]

# remove predators
traits<- traits[traits$spec_code!="NAB_OSC",]
traits<- traits[traits$spec_code!="COC_SP",]
traits<- traits[traits$spec_code!="PHI_SPU",]
#traits.both<-traits[traits$spec_code%in%colnames(hem.both),]    ### only need to do this for 2014 data? because used for fourth corner mostly?

# exclude species for which we don't have trait data. 
hem.both<-hem.both[ ,which( colnames(hem.both) %in% traits$spec_code )]
hem.2014<-hem.2014[ ,which( colnames(hem.2014) %in% traits$spec_code )]

traits<-traits[traits$spec_code%in%colnames(hem.2014),]
rownames(traits)<-traits$spec_code
traits<-traits[,-c(1)]

## There are NA's in the dataset. Remove.
leaftraits.full<-na.omit(leaftraits.full)

#select subset traits to prevent collinearity
leaftraits.full<-merge(leaftraits.full, env[,c(1,10)], by="vialID")
rownames(leaftraits.full)<-leaftraits.full$vialID
leaftraits.full<-leaftraits.full[,c(3,4,6,7,10,11)]
colnames(leaftraits.full)<-c("percN", "percP", "meanSLA", "meanWatercontent", "CN", "morphotype")

## Create leaftraits data (2014 only)
leaftraits<-leaftraits.full[rownames(leaftraits.full)%in%env.2014$vialID,]



# set different level of morphotype as intercept
leaftraits$morphotype<-factor(leaftraits$morphotype, levels=c("H", "G", "P"))

# load gps coordinates
setwd("/Users/elsketielens/phd/data/HDIM/branch clipping/data/data for R")
xy<-read.table("coordinatesHDIM.csv", header=TRUE, sep=",")
xy<-xy[xy$vialID%in%env$vialID,]
xy.2014<-xy[xy$vialID %in% env.2014$vialID,]
xy.both<-xy[xy$vialID %in% env.both$vialID,]
```
<br>

## Results

<br>
<br>

#### Some context: Richness across hybrid sites 
<br>
```{r describe_insect_community, echo=FALSE, warning=FALSE}
ord<-c( "ERO",  "KAIY", "OLAA", "ALI")

plot(ChaoRichness(as.data.frame(t(hem.both)), datatype = "abundance")[,2]~factor(env.both$site, as.ordered(ord)),  ylab="Estimated species richness (chao)", xlab="sites (ordered by substrate age)")
plot(exp(ChaoShannon(as.data.frame(t(hem.both)), datatype = "abundance")[,2])~factor(env.both$site, as.ordered(ord)),  ylab="Exp Estimated Shannon (hill 1)", xlab="sites (ordered by substrate age)")
plot(1/(1-ChaoSimpson(as.data.frame(t(hem.both)), datatype = "abundance")[,2])~factor(env.both$site, as.ordered(ord)),  ylab="1/1-Estimated Simpson (hill 2)", xlab="sites (ordered by substrate age)")

```

<br>

#### (i) How do herbivore communities respond to phenotypic variation?

<br>

```{r}

p<-metaMDS(hem.both, distance="bray", k=2, trymax=100)
ordiplot(p,type="n")
orditorp(p,display="species",col="red",air=0.01)
orditorp(p,display="sites",cex=1.25,air=0.01)

ordiplot(p,type="n")
orditorp(p,display="species",col="red",air=0.01)
orditorp(p,display="sites",pch=19, col="blue", label=F)
ordiellipse(p,groups=env.both$morphotype,draw="polygon",col="grey90",label=T, kind="se")
```
<br>

<br>

Are communities distinctly different between different morphotype trees within the same site?

<br>

```{r}
hem.ero<-hem[env$site=="ERO",]
hem.ali<-hem[env$site=="ALI",]
hem.kaiy<-hem[env$site=="KAIY",]
hem.olaa<-hem[env$site=="OLAA",]


#par(mfrow=c(2,2))

moment<-metaMDS(hem.ero, k=3)
mds.fig <- ordiplot(moment, type = "none", main="Escape road old flow")
text(mds.fig, "species", cex=0.5)
points(mds.fig, "sites", pch = 15, col = "blue", select = env[env$site=="ERO",]$morphotype == "H")
points(mds.fig, "sites", pch = 19, col = "red", select = env[env$site=="ERO",]$morphotype == "P")

Lab<-c( "Hybrid", "Pubescent")
color<-c("blue", "red")
legend(  "topright", legend=Lab,pch=c(15,19),pt.bg=c(color),pt.cex=1.1,cex=0.9,xpd=TRUE)



moment<-metaMDS(hem.kaiy, k=3)
mds.fig <- ordiplot(moment, type = "none", main="Kaiholena Young flow")
text(mds.fig, "species", cex=0.5)
points(mds.fig, "sites", pch = 15, col = "blue", select = env[env$site=="KAIY",]$morphotype == "G")
points(mds.fig, "sites", pch = 19, col = "red", select = env[env$site=="KAIY",]$morphotype == "P")

Lab<-c("Glabrous",  "Pubescent")
color<-c("blue", "red")
legend(  "topright", legend=Lab,pch=c(22,19),pt.bg=c(color),pt.cex=1.1,cex=0.9,xpd=TRUE)



moment<-metaMDS(hem.ali, k=3)
mds.fig <- ordiplot(moment, type = "none", main="Alili Springs")
text(mds.fig, "species",cex=0.5)
points(mds.fig, "sites", pch = 19, col = "green", select = env[env$site=="ALI",]$morphotype == "G")
points(mds.fig, "sites", pch = 15, col = "blue", select = env[env$site=="ALI",]$morphotype == "H")


Lab<-c("G", "H")
color<-c("green", "blue")
legend(  "topright", legend=Lab,pch=c(19,15),pt.bg=c(color),pt.cex=1.1,cex=0.9,xpd=TRUE)




moment<-metaMDS(hem.olaa, k=3)
mds.fig <- ordiplot(moment, type = "none", main="HAVO Olaa")
text(mds.fig, "species", cex=0.5)
points(mds.fig, "sites", pch = 19, col = "green", select = env[env$site=="OLAA",]$morphotype == "G")
points(mds.fig, "sites", pch = 15, col = "blue", select = env[env$site=="OLAA",]$morphotype == "H")

Lab<-c("Glabrous", "Hybrid")
color<-c("green", "blue" )
legend(  "topright", legend=Lab,pch=c(19,15),pt.bg=c(color),pt.cex=1.1,cex=0.9,xpd=TRUE)




adonis(hem.ali~env[env$site=="ALI",]$morphotype, by="margin")
adonis(hem.kaiy~env[env$site=="KAIY",]$morphotype, by="margin")
adonis(hem.ero~env[env$site=="ERO",]$morphotype, by="margin")
adonis(hem.olaa~env[env$site=="OLAA",]$morphotype, by="margin")




```

<br>

There are differences in species richness as well as species composition across sites. Diversity is highest at the high productivity sites midway the Big Island chronosequence. Species composition varies between sites, though many species are shared in common.  
<br>


#### (ii) Which plant traits account for herbivore community responses to host-plant phenotype?



<br>

```{r Cor}
# partition variance between environment and spatial
test<-rda(decostand(hem, "hel"), env$z.age, xy[,2:3])
test<-varpart(vegdist(hem), env$z.age, xy[,2:3])
capscale(vegdist(hem)~ xy[,2]+xy[,3])

# appears that spatial is not important at all across all 170 datapoints
# try again on subset with leaftraits
xy.sub<-xy[xy$vialID%in%rownames(leaftraits.full),]
test<-varpart(vegdist(decostand(hem.sub, "hel")), leaftraits.full[,c(1,2,3,4,6)], xy.sub[,2:3])
test
plot(test)
test<-rda(decostand(hem.2014, "hel"), leaftraits[,c(1,2,3,4,6)], xy.2014[,2:3])
test<-varpart(vegdist(decostand(hem.2014, "hel")), leaftraits[,c(1,2,3,4,6)], xy.2014[,2:3])
test.wout<-rda(decostand(hem.2014, "hel"), leaftraits[,c(1,2,3,4,6)], env.2014$site)
test.spat<-rda(decostand(hem.2014, "hel"),  xy.2014[,2:3])
anova(test)

adonis(hem.2014~leaftraits$percN+leaftraits$meanWatercontent+leaftraits$meanSLA+leaftraits$percP+xy.2014$long+env.2014$morphotype+xy.2014$lat, by="margin")


adonis(hem.2014~xy.2014$lat+xy.2014$long+leaftraits$percN+leaftraits$percP+leaftraits$meanSLA+leaftraits$meanWatercontent+env.2014$morphotype, by="margin")
# rda/varpart: when testing only 2014 hybrid zone sites from 2014, xy coordinates explain some. better explained still by leaftraits
# adonis seems to indicate latitude is important predictor- sig even after all other leaftrait variables
# while almost always sig (for 170 data points and 24), R2 is less than 0.1 in all cases  




adonis(hem.2014~leaftraits$morphotype+leaftraits$percN+leaftraits$meanWatercontent+leaftraits$meanSLA+leaftraits$percP+scale(xy.2014$lat,scale=T,center=T), by="margin")


adonis(hem.2014~env.2014$tree.height+leaftraits$morphotype+leaftraits$meanSLA+leaftraits$percN+env.2014$z.age+xy.2014$lat+xy.2014$long+env.2014$site, by="margin")


### try again for all 2014, not just hybrid

env.sub<-env[env$vialID%in%rownames(leaftraits.full),]
hem.sub<-hem[rownames(hem)%in%rownames(leaftraits.full),]

adonis(hem.sub~leaftraits.full$morphotype+leaftraits.full$meanSLA+leaftraits.full$percN+env.sub$z.age+env.sub$site, by="margin")









mod.leaf<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+leaftraits$meanWatercontent+leaftraits$percP)
mod.leaf.age<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+env.2014$age)
mod.leaf.tree<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+env.2014$tree.height)
mod.leaf.age.site<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+env.2014$tree.height+env.2014$z.age+env.2014$site+leaftraits$meanWatercontent+leaftraits$percP)
mod.leaf.site<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+env.2014$site+leaftraits$meanWatercontent+leaftraits$percP)


modfull<-cca(hem[rownames(hem)%in%rownames(leaftraits),]~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+env.2014$tree.height+env.2014$z.age+env.2014$site+leaftraits$meanWatercontent+leaftraits$percP)
m1<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+env.2014$tree.height+env.2014$z.age+env.2014$site+leaftraits$meanWatercontent+leaftraits$percP)
m2<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+leaftraits$percP+leaftraits$meanWatercontent)
m3<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+env.2014$tree.height+env.2014$site+leaftraits$meanWatercontent+leaftraits$percP)
m4<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+env.2014$z.age+env.2014$site+leaftraits$meanWatercontent+leaftraits$percP)
anova(modfull,m4)

m5<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+env.2014$site)
m6<-cca(hem.2014~leaftraits$morphotype+leaftraits$meanSLA+env.2014$site )       
  anova(m5, by = "margin", step=200 )      
  anova(m5,m6)
m7<-cca(hem.2014~leaftraits$morphotype+env.2014$site )   
anova(m6,m7)
m8<-cca(hem.2014~env.2014$site )  
anova(m7,m8)
# model with just site is best


# compare contribution of different predictors
mod.leaf
mod.leaf.age
mod.leaf.tree
mod.leaf.site
mod.leaf.age.site


plot(mod.leaf, type="none")
points(mod.leaf, display="sites")
text(mod.leaf, display="species")

anova(mod.leaf.age.site, by = "margin", step=200 )



ordiplot(mod.leaf)
ordiplot(mod.leaf.age)

# spread into hybrid versus pubescent/glabrous is explained by SLA, tree height Spread between pubescent and glabrous explained by age 




plot(mod.leaf.age.site, type="none")
points(mod.leaf.age.site, display="sites", pch=19)
#text(mod.leaf.age.site, display="species")
plot(envfit(mod.leaf.age.site, leaftraits[, c(1,2,3,4,6)]))



vare.cca <- cca(decostand(hem.2014, "total") ~ ., data = leaftraits[,c(1,6)]) 

plot(vare.cca, display = "sites", scaling = 3, type="points")
text(vare.cca, scaling = 3, display = "bp", pch=4) 
text(vare.cca, display = "species", scaling = 3, col="red", cex=0.5)



leaftraits.new<-leaftraits[leaftraits$vialID%in%env.yng$vialID,]
leaftraits.new<-merge(leaftraits.new,env.yng, by="vialID")
plot(m9)
m2
```


#### Do CCA with data lumped for 2014 and 2015
```{r}
hem.all<-hem
hem.all$vialID<-rownames(hem.all)
hem.all<-merge(hem.all, env, by="vialID")
hem.all<-hem.all[,-c(87:97)]
melt<-melt(hem.all)
hem.all<-dcast(melt, site+plot~variable, sum)

env.all<-hem.all[,c(1,2)]
hem.all$ID<-paste(hem.all$site, hem.all$plot, sep="_")
rownames(hem.all)<-hem.all$ID


env.alt<-env[env$vialID%in%rownames(leaftraits.full),]
#env.alt<-env.alt[,-1]
env.alt<-merge(env.alt,env.all, by=c("site","plot"))
env.alt$ID<-paste(env.alt$site, env.alt$plot, sep="_")
env.alt$DW[4]<-env.alt$DW[4]+env.alt$DW[5]
env.alt<-env.alt[-5,]
env.alt$DW[71]<-env.alt$DW[71]+env.alt$DW[72]
env.alt<-env.alt[-72,]
hem.all<-hem.all[hem.all$ID %in% env.alt$ID,]
env.alt<-env.alt[env.alt$ID %in% hem.all$ID,]

leaftraits.alt<-leaftraits.full[rownames(leaftraits.full)%in%env.alt$vialID,]

hem.all<-hem.all[,-c(1,2,86)]
colSums(hem.all)
rowSums(hem.all)


# data looks good- use hem.all for merged 2014 and 2015 data






#######################################






plot(mod.leaf.age.site, type="none")
points(mod.leaf.age.site, display="sites", pch=19)
#text(mod.leaf.age.site, display="species")
plot(envfit(mod.leaf.age.site, leaftraits[, c(1,2,3,4,6)]))



vare.cca <- cca(decostand(hem.all, "total") ~ ., data = leaftraits.alt) 

plot(vare.cca, display = "sites", scaling = 3, type="points")
text(vare.cca, scaling = 3, display = "bp", pch=4) 
text(vare.cca, display = "species", scaling = 3, col="red", cex=0.5)
anova(vare.cca, by = "margin", step=200 )

adonis(hem.all~leaftraits.alt$morphotype+leaftraits.alt$meanSLA+leaftraits.alt$percN+env.alt$z.age, by="margin")


```




<br>

#### (iii) Which insect traits account for herbivore community responses to host-plant phenotype
<br>

<br>

Are communities distinctly different between different morphotype trees across all youngish sites?
```{r}

moment<-metaMDS(hem.both, k=3)
mds.fig <- ordiplot(moment, type = "none", main="all hybrid zone sites")
#text(mds.fig, "species")
points(mds.fig, "sites", pch = 19, col = "green", select = env.both$morphotype == "G")
points(mds.fig, "sites", pch = 15, col = "blue", select = env.both$morphotype == "H")
points(mds.fig, "sites", pch = 17, col = "red", select = env.both$morphotype == "P")

Lab<-c("Glabrous", "Hybrid", "Pubescent")
color<-c("green", "blue", "red")
legend(  "topright", legend=Lab,pch=c(19,15,17),pt.bg=c(color),pt.cex=1.1,cex=0.9,xpd=TRUE)

## across all sites (n=170)
moment<-metaMDS(hem, k=3)
mds.fig <- ordiplot(moment, type = "none", main="all sites")
text(mds.fig, "species",cex=0.5)
points(mds.fig, "sites", pch = 19, col = "grey", select = env$morphotype == "G")
points(mds.fig, "sites", pch = 15, col = "blue", select = env$morphotype == "H")
points(mds.fig, "sites", pch = 17, col = "red", select = env$morphotype == "P")

Lab<-c("Glabrous", "Hybrid", "Pubescent")
color<-c("grey", "blue", "red")
legend(  "topright", legend=Lab,pch=c(19,15,17),pt.bg=c(color),pt.cex=1.1,cex=0.9,xpd=TRUE)

# for all sites without double measures (n=97)
moment<-metaMDS(hem.all, k=3)
mds.fig <- ordiplot(moment, type = "none", main="all sites")
text(mds.fig, "species",cex=0.5)
points(mds.fig, "sites", pch = 19, col = "grey", select = env$morphotype == "G")
points(mds.fig, "sites", pch = 15, col = "blue", select = env$morphotype == "H")
points(mds.fig, "sites", pch = 17, col = "red", select = env$morphotype == "P")

```
<br>




#

adonis(hem.2014 ~  leaftraits$percN +leaftraits$percP+leaftraits$morphotype  + leaftraits$meanWatercontent +leaftraits$meanSLA +env.2014$site , by="terms")

# phosphorus and morphotype are only sig when only/first term
# mean SLA and mean water content are always sig no matter their placement
# perc N significant unless after SLA and before water content



<br>
What does diversity and abundance look like across these sites with multiple phenotypes?
```{r}
plot(1/(1-ChaoSimpson(as.data.frame(t(hem.ero)) )[,2])~env[env$site=="ERO",]$morphotype)
plot(1/(1-ChaoSimpson(as.data.frame(t(hem.kaiy)) )[,2])~env[env$site=="KAIY",]$morphotype)
plot(1/(1-ChaoSimpson(as.data.frame(t(hem.olaa)) )[,2])~env[env$site=="OLAA",]$morphotype)
plot(1/(1-ChaoSimpson(as.data.frame(t(hem.ali)) )[,2])~env[env$site=="ALI",]$morphotype)

plot(ChaoRichness(as.data.frame(t(hem.ero)) )[,2]~env[env$site=="ERO",]$morphotype)
plot(ChaoRichness(as.data.frame(t(hem.kaiy)) )[,2]~env[env$site=="KAIY",]$morphotype)
plot(ChaoRichness(as.data.frame(t(hem.olaa)) )[,2]~env[env$site=="OLAA",]$morphotype)
plot(ChaoRichness(as.data.frame(t(hem.ali)) )[,2]~env[env$site=="ALI",]$morphotype)

plot(rowSums(hem.ero)~env[env$site=="ERO",]$morphotype)
plot(rowSums(hem.kaiy)~env[env$site=="KAIY",]$morphotype)
plot(rowSums(hem.olaa)~env[env$site=="OLAA",]$morphotype)
plot(rowSums(hem.ali)~env[env$site=="ALI",]$morphotype)

anova(lm(1/(1-ChaoSimpson(as.data.frame(t(hem.ero)) )[,2])~env[env$site=="ERO",]$morphotype))
#anova(lm(1/(1-ChaoSimpson(as.data.frame(t(hem.kaiy)) )[,2])~env[env$site=="KAIY",]$morphotype))
anova(lm(1/(1-ChaoSimpson(as.data.frame(t(hem.olaa)) )[,2])~env[env$site=="OLAA",]$morphotype))
anova(lm(1/(1-ChaoSimpson(as.data.frame(t(hem.ali)) )[,2])~env[env$site=="ALI",]$morphotype))

anova(lm(rowSums(hem.kaiy)~env[env$site=="KAIY",]$morphotype))
# p<0.05
anova(lm(rowSums(hem.olaa)~env[env$site=="OLAA",]$morphotype))
# p<0.1


# Across all 12 sites:
env.both$morphotype<-factor(env.both$morphotype, levels=c("H", "G", "P"))
plot(rowSums(hem.both)~env.both$morphotype, ylim=c(0,300), ylab="Abundance")
anova(lm(rowSums(hem.both)~env.both$morphotype))
TukeyHSD(aov(rowSums(hem.both)~env.both$morphotype))
# p<0.05

tem<-1-ChaoSimpson(as.data.frame(t(hem.both), na.rm=TRUE) )
#tem[10,2]<-1
#tem[65,2]<-1
tem<-ifelse(tem[,2]==0,0,1/tem[,2])
anova(lm(tem~env.both$morphotype))
anova(lmer(tem~env.both$morphotype+(1|env.both$site)))

tem<-exp(ChaoShannon(as.data.frame(t(hem.both)) )[,2])
#tem[2]<-1
#tem[26]<-1 
anova(lm(tem~env.both$morphotype))
anova(lmer(tem~env.both$morphotype+(1|env.both$site)))

tem<-ChaoRichness(as.data.frame(t(hem.both)))[,2]
anova(lm(tem~env.both$morphotype))
anova(lmer(tem~env.both$morphotype+(1|env.both$site)))
plot(tem~env.both$morphotype, ylab="Species Richness")

# richness does diff sig between morphotype, diversity shannon or simpson does not.
# ie the difference between glabrous and hybrid and glabrous and pubescent is in rare species- glabrous has more rare species than the other two morphotypes, but once you weigh rare species less, there are no differences in diversity



### plots for all sites
plot(diversity(hem)/log(specnumber(hem))~env$morphotype, ylab="Pielou's Evenness")
plot(rowSums(hem)~env$morphotype,  ylab="Abundance",xlab="Tree morphotype", main="across big island (n=170)", ylim=c(0,350))

tem<-exp(ChaoShannon(as.data.frame(t(hem)) )[,2])
plot(tem~env$morphotype, ylab="Exp Shannon Index",main="across big island (n=170)", xlab="Tree morphotype")
```

<br>


```{r}
#mod.leaf.site<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+leaftraits$meanWatercontent+leaftraits$percP+env.2014$site, na.rm=TRUE)


#mod.leaf.site
# 56% of variation constrained

#mod.leaf<-cca(hem.2014~leaftraits$percN+leaftraits$morphotype+leaftraits$meanSLA+leaftraits$meanWatercontent+leaftraits$percP)
mod.leaf


#plot(mod.leaf.site, type="none")
#points(mod.leaf.site, display="sites")
#text(mod.leaf.site, display="species")
#plot(envfit(mod.leaf.site, temp5[,c(3,4,5,6,8)]))

#anova(mod.leaf.site, by = "margin", step=200 )


```

How much variation is explained by spatial proximity?
```{r}





```

<br>


```{r fourth1}
library(mvabund)


traits$feeding_guild<-factor(traits$feeding_guild)
leaftraits<-leaftraits[order(rownames(leaftraits)),]
hem.2014<-hem.2014[order(rownames(hem.2014)),]
hem.2014<-hem.2014[,colnames(hem.2014)%in%rownames(traits)]
traits<-traits[order(rownames(traits)),]

mod<-manyglm(mvabund(hem.2014)~leaftraits$percN+leaftraits$percP+leaftraits$morphotype+leaftraits$meanWatercontent+leaftraits$meanSLA, family="negative binomial")

library(lattice)
a        = max( abs(mod$coefficients) )
colort   = colorRampPalette(c("blue","white","red")) 
plot.spp = levelplot(as.matrix(mod$coefficients), xlab="Environmental Variables",
 ylab="Species traits", col.regions=colort(100), at=seq(-a, a, length=100),
 scales = list( x= list(rot = 45)))

plot.spp


```

#### 2d) Test strength of correlation between insect traits and host plant characteristics?

<br>
<br>

```{r fourth2}
ftspp1=traitglm(mvabund(hem.2014),leaftraits,method="glm1path", family="poisson")

library(lattice)
a        = max( abs(ftspp1$fourth.corner) )
colort   = colorRampPalette(c("blue","white","red")) 
plot.spp = levelplot(t(as.matrix(ftspp1$fourth.corner)), xlab="Environmental Variables",
 ylab="Species traits", col.regions=colort(100), at=seq(-a, a, length=100),
 scales = list( x= list(rot = 45)))

print(plot.spp)




#levels(traits$polyphagy)<-c(levels(traits$polyphagy) ,"zgeneralist")
#traits$polyphagy[traits$polyphagy=="generalist"]<-"zgeneralist"
#traits$polyphagy<-factor(traits$polyphagy)
#levels(traits$polyphagy)

######################


hem.sub<-hem.2014[,colnames(hem.2014)%in%rownames(traits)]
traits<-traits[rownames(traits)%in%colnames(hem.sub),]
traits<-traits[order(rownames(traits)),]
hem.sub<-hem.sub[,order(colnames(hem.sub))]
# Run 4th corner model
model4th<-traitglm(hem.2014, leaftraits[,c(1,2,3,4,6)], traits,  method="manyglm", family="negative.binomial")

# call coefficients
model4th$fourth

# plot coefficients
a        = max( abs(model4th$fourth.corner),na.rm=T )
colort   = colorRampPalette(c("blue","white","red")) 
plot.4th = levelplot(t(as.matrix(model4th$fourth.corner)), xlab="Host Plant Traits",
              ylab="Species traits", col.regions=colort(100), at=seq(-a, a, length=100),
                     scales = list( x= list(rot = 45), fontface="bold"))


print(plot.4th)

colnames(leaftraits)<-c("% Nitrogen", "% Phosphorus", "trichomes -", "SLA", "watercontent", "C:N")
levels(leaftraits$`trichomes -`)<-c(levels(leaftraits$`trichomes -`), "none","intermediate", "high")
leaftraits$`trichomes -`[leaftraits$`trichomes -`=="G"]<- "none"
leaftraits$`trichomes -`[leaftraits$`trichomes -`=="H"]<- "intermediate"
leaftraits$`trichomes -`[leaftraits$`trichomes -`=="P"]<- "high"
leaftraits$`trichomes -`<-factor(leaftraits$`trichomes -`)


colnames(traits)<-c("length", "feeding guild - ", "location nymph - ", "hosts - ")
levels(traits$`hosts:`)<-c(levels(traits$`hosts:`), "generalist","specialist")
traits$`hosts:`[traits$`hosts:`=="zgeneralist"]<-"generalist"
traits$`hosts:`<-factor(traits$`hosts:`)
traits$`hosts:`<-factor(traits$`hosts:`,levels=c("specialist", "generalist"))




######################



ftSmall=traitglm(hem.2014,leaftraits[,c(1,2,3,4,6)],traits)
anova.traitglm(ftSmall, nBoot=999, show.time=T)
# does permutations = work for this function?

model4th<-traitglm(hem.sub, leaftraits, traits,  method="glm1path", family="negative.binomial")
model4th$fourth

```





```{r}
hem.both$vialID<-rownames(hem.both)
hem.both.long<-melt(hem.both)
hem.both.long<-merge(env.both, hem.both.long, by="vialID")

hem.2014$vialID<-rownames(hem.2014)
leaftraits$vialID<-rownames(leaftraits)
hem.2014.long<-melt(hem.2014)
hem.2014.long<-merge(leaftraits, hem.2014.long, by="vialID")
hem.2014.long<-merge(env[,1:2], hem.2014.long, by="vialID")


mod.season    =glmer(value    ~morphotype +perc   +(1+morphotype|variable)+(1|site),  data    =hem.both.long,   family    =poisson)
dotplot(ranef(mod.season,    condVar    =T))
summary(mod.season)

mod.no.season    =glmer(value    ~morphotype    +(1|variable),    data    =hem.both.long,family  =poisson)
anova(mod.season,    mod.no.season)

hem.both.long$percN.standard<-(hem.both.long$percN-mean(hem.both.long$percN))/sd(hem.both.long$percN)
hem.both.long$percP.standard<-(hem.both.long$percP-mean(hem.both.long$percP))/sd(hem.both.long$percP)
hem.both.long$meanSLA.standard<-(hem.both.long$meanSLA-mean(hem.both.long$meanSLA))/sd(hem.both.long$meanSLA)
hem.both.long$meanWatercontent.standard<-(hem.both.long$meanWatercontent-mean(hem.both.long$meanWatercontent))/sd(hem.both.long$meanWatercontent)








############## 2014 only


hem.2014.long$percN.standard<-(hem.2014.long$percN-mean(hem.2014.long$percN))/sd(hem.2014.long$percN)
hem.2014.long$percP.standard<-(hem.2014.long$percP-mean(hem.2014.long$percP))/sd(hem.2014.long$percP)
hem.2014.long$meanSLA.standard<-(hem.2014.long$meanSLA-mean(hem.2014.long$meanSLA))/sd(hem.2014.long$meanSLA)
hem.2014.long$meanWatercontent.standard<-(hem.2014.long$meanWatercontent-mean(hem.2014.long$meanWatercontent))/sd(hem.2014.long$meanWatercontent)


# not enough datapoints for all predictors, only include half
mod.leaf    =glmer(value    ~morphotype +percN.standard+meanSLA.standard   +(1+morphotype+percN.standard+meanSLA.standard|variable),  data    =hem.2014.long,   family    =poisson)
dotplot(ranef(mod.leaf,    condVar    =T))
summary(mod.season)

mod.no.leaf    =glmer(value    ~morphotype   +percN.standard+meanSLA.standard +(1|variable),    data    =hem.2014.long,family  =poisson)
anova(mod.leaf,    mod.no.leaf)
# model with species specific intercept is better


# check for other half predictors
mod.leaf.2    =glmer(value    ~percP.standard+meanWatercontent.standard   +(1+percP.standard+meanWatercontent.standard|variable),  data    =hem.2014.long,   family    =poisson)
summary(mod.leaf.2)

anova(mod.leaf.2,    mod.leaf,mod.no.leaf)

dotplot(ranef(mod.leaf.2,    condVar    =T))

mod.leaf.2    =glmer(value    ~percP.standard+meanWatercontent.standard   +(1+percP.standard+meanWatercontent.standard|variable),  data    =hem.2014.long,   family    =poisson)
summary(mod.leaf.2)


mod1    =glmer(value    ~percN.standard+meanWatercontent.standard +morphotype  +(1+percN.standard+meanWatercontent.standard+morphotype|variable),  data    =hem.2014.long,   family    =poisson)
anova(mod.leaf,mod.leaf.2,mod1)
mod2    =glmer(value    ~percP.standard+meanWatercontent.standard + percN.standard+meanSLA.standard +(1+percP.standard+meanWatercontent.standard+percN.standard+meanSLA.standard|variable),  data    =hem.2014.long,   family    =poisson)
anova(mod1,mod2,mod.leaf.2)
# keep morphotype

# model without N, with other predictors
mod3    =glmer(value    ~percP.standard+meanWatercontent.standard +morphotype +meanSLA.standard +(1+percP.standard+meanWatercontent.standard+morphotype+meanSLA.standard|variable),  data    =hem.2014.long,   family    =poisson)

# full model:
mod4    =glmer(value    ~percP.standard+meanWatercontent.standard +morphotype +meanSLA.standard+percN.standard +(1+percP.standard+meanWatercontent.standard+morphotype+meanSLA.standard+percN.standard|variable),  data    =hem.2014.long,   family    =poisson)
anova(mod3,mod1,mod4)
# full model is best

mod5    =glmer(value    ~percP.standard +morphotype +meanSLA.standard+percN.standard +(1+percP.standard+morphotype+meanSLA.standard+percN.standard|variable),  data    =hem.2014.long,   family    =poisson)
mod6    =glmer(value    ~meanWatercontent.standard +morphotype +meanSLA.standard+percN.standard +(1+meanWatercontent.standard+morphotype+meanSLA.standard+percN.standard|variable),  data    =hem.2014.long,   family    =poisson)
mod7    =glmer(value    ~meanWatercontent.standard +morphotype +percN.standard +(1+meanWatercontent.standard+morphotype+percN.standard|variable),  data    =hem.2014.long,   family    =poisson)
mod8    =glmer(value    ~percP.standard+meanWatercontent.standard  +meanSLA.standard+percN.standard +(1+percP.standard+meanWatercontent.standard+meanSLA.standard+percN.standard|variable),  data    =hem.2014.long,   family    =poisson)

anova(mod4,mod5,mod6,mod3,mod7,mod8)
# mod4 is best- ie full model. 

mod9    =glmer(value    ~percP.standard+morphotype +(1+percP.standard+morphotype|variable),  data    =hem.2014.long,   family    =poisson)

summary(mod4)
# important predictors are phorphorus and morphotype

## Full model is best, sig predictors in full model are phosphorus and morphotype. 
# species specific intercepts explain much better
#


dotplot(ranef(mod4,condVar=T))
# look at dotplot on a leaftrait by leaftrait basis

mod    =glmer(value    ~percP.standard +(1+percP.standard|variable),  data    =hem.2014.long,   family    =poisson)
dotplot(ranef(mod,condVar=T))
mod    =glmer(value    ~percN.standard +(1+percN.standard|variable),  data    =hem.2014.long,   family    =poisson)
dotplot(ranef(mod,condVar=T))
mod    =glmer(value    ~meanSLA.standard +(1+meanSLA.standard|variable),  data    =hem.2014.long,   family    =poisson)
dotplot(ranef(mod,condVar=T))
mod    =glmer(value    ~meanWatercontent.standard +(1+meanWatercontent.standard|variable),  data    =hem.2014.long,   family    =poisson)
dotplot(ranef(mod,condVar=T))
mod    =glmer(value    ~morphotype +(1+morphotype|variable),  data    =hem.2014.long,   family    =poisson)
dotplot(ranef(mod,condVar=T))



hemlong<-hem.2014.long[hem.2014.long$variable==c("GRE_PSI"),]
hemlong<-rbind(hemlong,hem.2014.long[hem.2014.long$variable==c("OCE_VUL"),])
hemlong<-rbind(hemlong,hem.2014.long[hem.2014.long$variable==c("OPU_SP"),])
hemlong<-rbind(hemlong,hem.2014.long[hem.2014.long$variable==c("LEI_SP"),])
hemlong<-rbind(hemlong,hem.2014.long[hem.2014.long$variable==c("ORT_MET"),])
hemlong<-rbind(hemlong,hem.2014.long[hem.2014.long$variable==c("KOA_HAW"),])
hemlong<-rbind(hemlong,hem.2014.long[hem.2014.long$variable==c("SAR_ADO"),])
hemlong<-rbind(hemlong,hem.2014.long[hem.2014.long$variable==c("GEO_SP"),])
hemlong<-rbind(hemlong,hem.2014.long[hem.2014.long$variable==c("PAR_PYR"),])
hemlong<-rbind(hemlong,hem.2014.long[hem.2014.long$variable==c("TYM_TYM"),])


cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000", "green")
  
  
  ggplot(data=hemlong, aes(x=percN.standard, y=log(value), group=variable, colour=variable)) +
#    geom_point()+
  geom_smooth(method=lm, se=F,level=0.95)+
  ylab("Abundance")+ xlab("standardized foliar nitrogen %")+
  scale_colour_manual(values=cbPalette, name="Species",  labels=c("GRE_PSI Aphididae", "OCE_VUL Lygaeidae", "OPU_SP Miridae", "LEI_SP Delphacidae", "ORT_MET Miridae", "KOA_HAW Miridae", "SAR_ADO Miridae", "GEO_SP Geometridae (Lepidoptera)", "PAR_PYR Triozidae", "TYM_TYM Pseudococcidae"))+
  theme_bw()


ggplot(data=hemlong, aes(x=percP.standard, y=log(value), group=variable, colour=variable)) +
 #   geom_point()+
  geom_smooth(method=lm, se=F,level=0.95)+
    ylab("Abundance")+ xlab("standardized foliar phosphorus %")+
  scale_colour_manual(values=cbPalette, name="Species",  labels=c("GRE_PSI Aphididae", "OCE_VUL Lygaeidae", "OPU_SP Miridae", "LEI_SP Delphacidae", "ORT_MET Miridae", "KOA_HAW Miridae", "SAR_ADO Miridae", "GEO_SP Geometridae (Lepidoptera)", "PAR_PYR Triozidae", "TYM_TYM Pseudococcidae"))+
  theme_bw()

ggplot(data=hemlong, aes(x=meanSLA.standard, y=log(value), group=variable, colour=variable)) +
#    geom_point()+
  geom_smooth(method=lm, se=F,level=0.95)+
    ylab("Abundance")+ xlab("standardized mean specific leaf area")+
  scale_colour_manual(values=cbPalette, name="Species",  labels=c("GRE_PSI Aphididae", "OCE_VUL Lygaeidae", "OPU_SP Miridae", "LEI_SP Delphacidae", "ORT_MET Miridae", "KOA_HAW Miridae", "SAR_ADO Miridae", "GEO_SP Geometridae (Lepidoptera)", "PAR_PYR Triozidae", "TYM_TYM Pseudococcidae"))+
  theme_bw()

ggplot(data=hemlong, aes(x=meanWatercontent.standard, y=log(value), group=variable, colour=variable)) +
#    geom_point()+
  geom_smooth(method=lm, se=F,level=0.95)+
    ylab("Abundance")+ xlab("standardized mean water content")+
  scale_colour_manual(values=cbPalette, name="Species",  labels=c("GRE_PSI Aphididae", "OCE_VUL Lygaeidae", "OPU_SP Miridae", "LEI_SP Delphacidae", "ORT_MET Miridae", "KOA_HAW Miridae", "SAR_ADO Miridae", "GEO_SP Geometridae (Lepidoptera)", "PAR_PYR Triozidae", "TYM_TYM Pseudococcidae"))+
  theme_bw()




#
```





## Conclusions
<br>
<br>

##### (i) How do herbivore communities respond to phenotypic variation?
There is variation in species composition across sites and metrosideros traits. NMDS plots indicate differences between sites and volcanoes in species composition. 

##### (ii) Which plant traits account for herbivore community responses to host-plant phenotype?
CCA results point to the importance of foliar nitrogen, leaf morphotype and specific leaf area. However, there is still a large part of variation not explained. 

##### (iii) Which insect traits account for herbivore community responses to host-plant phenotype
Community composition varies across metrosideros traits. Feeding guild is correlated with foliar %N, where leaf chewers (sig) and sap feeders (NS) are more abundant on higher nutrient trees. Feeding guild is correlated with morphotype, where high trichome density morphotype is correlated with higher abundance gallers and seed feeders, and lower abundances of chewers and sap feeders. Morphotype is positively correlated with proportion specialists (pubescent leaves show relatively higher abundances of specialists). Location of the nymph varies little across morphotypes, 



<br>




```{r}
temp<-hem.both
temp<-ifelse(hem.both>0,1,0)
beta.multi(temp, index.family="sor")

temp<-hem.olaa
temp<-ifelse(hem.olaa>0,1,0)
beta.multi(temp, index.family="sor")

beta.pair(temp)

beta.ceram.s<-beta.sample(temp, index.family="sor", sites=10, samples=100)








# NOT RUN {
# Read the data for Northern and Southern European cerambycids
ceram.s<-ifelse(hem.kaiy>0,1,0)
ceram.n<-ifelse(hem.ero>0,1,0)

# Resample 100 times the multiple-site dissimilarities
# for 10 countries.
beta.ceram.s<-beta.sample(ceram.s, index.family="sor", sites=10, samples=100)
beta.ceram.n<-beta.sample(ceram.n, index.family="sor", sites=10, samples=100)

# Plot the distributions of beta.SIM in Southern Europe (red) 
# and Northern Europe (blue)
plot(density(beta.ceram.s$sampled.values$beta.SIM), col="red", xlim=c(0,1))
lines(density(beta.ceram.n$sampled.values$beta.SIM), col="blue")

# Compute the p-value of difference in beta.SIM between South and North 
# (i.e. the probability of finding in the North a higher value than 
# in the South)
p.value.beta.SIM<-length(which(beta.ceram.s$sampled.values$beta.SIM<
beta.ceram.n$sampled.values$beta.SIM))/100

p.value.beta.SIM
# The result is 0 and we used 100 samples, so p<0.01
# }
```




